---
editor: visual
format: html
execute: 
  echo: false
  warning: false
  error: false
---

# Solo Project {-}

For this solo project I wanted to explore my home state of Washington.
It is a large state both in population and geography providing the potential for interesting comparisons and differentiation.
Education data in Washington State is also incredibly prevalent and well characterized.
Thus I explored some of the different state and district websites.

```{r}
library(tidyverse)
library(sf)
library(mosaic)

boundaries <- read_sf("C:\\Users\\ashry\\OneDrive\\Documents\\GitHub\\portfolio-AshERyan21\\bw\\data\\Washington_School_District_Boundaries.shp")
data <- read_csv("C:\\Users\\ashry\\OneDrive\\Documents\\GitHub\\portfolio-AshERyan21\\bw\\data\\Report_Card_WaKids_2023-24_School_Year_20250418.csv")

```

I found a data set detailing the school districts in Washington as well as a data set detailing kindergarten readiness scores by district, demographic, and topic.

```{r}
dim(data)
head(data, 3)
tail(data, 3)
```

The data includes a lot of information and metrics regarding the different metrics of readiness, as well as subcategory breakdowns.

```{r}
Data_Totals <- data |>
  filter(Measure == "NumberofDomainsReadyforKindergarten") |>
  group_by(DistrictName)|>
  summarize(total_ready = sum(Percent, na.rm = TRUE))

```

For my exploration I only really wanted to look at the number of domains ready for Kindergarten, and look at the sums of those percentages.
Therefor, I filtered by data to only include those in, "Number of Domains Ready for Kindergarten."

```{r}
#in boundaries its LEAName, Data_Totals its DistrictName
FullDATA <- boundaries |>
  left_join(Data_Totals, join_by(LEAName == DistrictName))
```

Then I joined the data with the necessary boundaries.

```{r}
LOGFullDATA <- FullDATA |>
  mutate(PTR = log(total_ready))
```


```{r}
#| fig-height: 10
#| fig-width: 10
#| fig-cap: >
#|  The percent total ready for Kindergarten across all categories with yellow representing the highest and purple the lowest. The grey areas are NA.   
#|  Created by Ashlyn R, Macalester College Introduction to Data Science, Spring 2025. Data from Washington Department of Education.
ggplot() +
  geom_sf(data = LOGFullDATA, aes(fill= total_ready), color = "white") +
  scale_fill_viridis_c(option = "viridis", direction = 1) +
   labs(title = "Washington State By County and Kindergarten Readiness", fill = "Percent Total Ready") +
  theme_map()
```

The state map above represents the total percent of all categories combined by state for readiness for kindergarten.
We see in this graph that there is a very large range of readiness.
The hotspot areas appear to be spread throughout the state; however, as a resident of the state I know that those are the areas with cities/ larger population centers.

Lastly, due to the range within the data I created a log scale of the percent total ready, potentially allowing for a better range analysis.
```{r}
#| fig-height: 10
#| fig-width: 10
#| fig-cap: >
#|   The percent total ready for Kindergarten across all categories with yellow representing the highest and purple the lowest. The grey areas are NA.
#|   Created by Ashlyn R, Macalester College Introduction to Data Science, Spring 2025. Data from Washington Department of Education.

ggplot() +
  geom_sf(data = LOGFullDATA, aes(fill= PTR), color = "white") +
  scale_fill_viridis_c(option = "viridis", direction = 1) +
  labs(title = "Washington State By County and Kindergarten Readiness (Logged)", fill = "Percent Total Ready (Logged)") +
  theme_map()
```

By putting the total ready on a log scale, the hotspots remain highlighted, however, the variation seen throughout the rest of the state is easier to be seen.
I think that this shows that throughout Washington readiness for kindergarted really varies.
The downside of the log scale map is that more areas do not have data: are grey, as those are the areas either with zeros/ and, or NAS.
This does takeaway some of what can be learned from the data.
