---
number-sections: true
execute: 
  echo: false
  warning: false
  error: false
fig-env: 'figure'
fig-pos: 'h'
fig-align: center
code-fold: false
---

# Exam 2 {-}

```{r}
library(tidytuesdayR)
library(tidyverse)
library(rnaturalearth)
library(sf)
```

```{r}
tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption
```

```{r}
str(fc)
```

```{r}
head(fc, 22)
```


```{r}
fc|>
  distinct(food_category)
```
```{r}
fcc <- fc |>
  mutate(food_category = fct_recode(food_category, 
                                    "Lamb" = "Lamb & Goat",
                                    "Milk" = "Milk - inc. cheese",
                                    "Wheat" = "Wheat and Wheat Products",
                                    "Nuts" = "Nuts inc. Peanut Butter"))
```

```{r}
fcc |>
  distinct(food_category)
```

Research Question 1: Which 5 countries consume the most food are?
(Show the answer as a 2-column table and as a graph, sort based on consumption)

```{r}
tc <- fcc |>
  group_by(country) |>
  summarize(total = sum(consumption)) |>
  arrange(desc(total)) |>
  head(5)
```


```{r}
tc
```
```{r}
ggplot(tc, aes(x = country, y = total)) +
  geom_col()
```

Research Question 2: Which top 5 countries consume each food are?
```{r}
fcc2 <- fcc |>
  select(country, food_category, consumption) |>
  group_by(food_category) |>
  slice_max(consumption, n= 5)
  
fcc2
```

```{r}
#| fig-height: 65
#| fig-width: 15
fcc2|>
  ggplot(aes(x = consumption, y = country)) +
  geom_col() +
  facet_wrap(~ food_category, ncol = 1)
```
Research Question 3: What does the consumption of each food look like?

Viz: Attempt 1
```{r}
ne_countries(returnclass = "sf") |>
  select(name, geometry) |>
  left_join(fcc |> select(-co2_emmission),
            join_by(name == country)) |>
  ggplot()+
  geom_sf(aes(fill = consumption)) +
  facet_wrap(~food_category) +
  theme(legend.position= "bottom")
```
Countries with Missing Consumption Data
```{r}

```


Viz: Attempt 2
```{r}
#| fig-height: 15
#| fig-width: 15
ne_countries(returnclass = "sf") |>
  select(name, geometry) |>
  mutate(name = ifelse(name == "United States of America", "USA", name)) |>
  mutate(name = ifelse(name == "Bosnia and Herz.", "Bosnia and Herzegovina", name)) |>
  mutate(name = ifelse(name == "Czechia", "Czech Republic", name)) |>
  mutate(name = ifelse(name == "Taiwan", "Taiwan. ROC", name)) |>
  left_join(fcc |> select(-co2_emmission),
            join_by(name == country)) |>
  pivot_wider(names_from = food_category,
              values_from = consumption) |>
  select(-"NA") |>
  pivot_longer(cols = c(-name, -geometry),
               names_to = "food_category",
               values_to = "consumption") |>
  ggplot()+
  geom_sf(aes(fill = consumption)) +
  facet_wrap(~food_category) +
  theme(legend.position = "bottom")

```
Food Consumption Statistics
```{r}
#| fig-height: 15
#| fig-width: 15
ne_countries(returnclass = "sf") |>
  select(name, geometry) |>
  mutate(name = ifelse(name == "United States of America", "USA", name)) |>
  mutate(name = ifelse(name == "Bosnia and Herz.", "Bosnia and Herzegovina", name)) |>
  mutate(name = ifelse(name == "Czechia", "Czech Republic", name)) |>
  mutate(name = ifelse(name == "Taiwan", "Taiwan. ROC", name)) |>
  left_join(
    fcc |> 
      select(-co2_emmission) |>
      group_by(food_category) |>
      mutate(consumption = (consumption - mean(consumption))/sd(consumption)),
    join_by(name == country)) |>
  pivot_wider(names_from = food_category,
              values_from = consumption) |>
  select(-"NA") |>
  pivot_longer(cols = c(-name, -geometry),
               names_to = "food_category",
               values_to = "consumption") |>
  ggplot()+
  geom_sf(aes(fill = consumption)) +
  facet_wrap(~food_category) +
  theme(legend.position = "bottom")

```





