---
title: "Exam 1"
number-sections: true
execute: 
  warning: true
fig-env: 'figure'
fig-pos: 'h'
fig-align: center
code-fold: false
---


```{r}
library(tidytuesdayR)
library(tidyverse)
```


```{r}
tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption
```

Grand Research Question: What does the consumption of each food category in each country look like?


```{r}
str(fc)
dim(fc)
```

```{r}
head(fc,22)
tail(fc,22)
```

We have three variables included in the grand research question, before we understand the big picture question we need to look at each one individually.

```{r}
ggplot(fc, aes(x = consumption)) + 
  geom_density(alpha = 0.5) + 
  labs(x = "Consumption (kg/person/year)") + 
  theme_classic()    
```

The graph above details the density of consumption, showcasing that consumption varies greatly but spikes closer to zero but with some outliers. 

```{r}
ggplot(fc, aes(x = food_category)) +
  geom_bar()
```


Understanding overall food consumption and each food category and each country.

```{r}
ggplot(fc, aes(y = consumption, x = food_category)) +
  geom_boxplot()
```

```{r}
#| fig-height: 22
#| fig-width: 11

ggplot(fc, aes(y = consumption, x = food_category)) +
  geom_col()

```


```{r}
#| fig-height: 22
#| fig-width: 11

ggplot(fc, aes(x = consumption)) + 
  geom_density() +
  labs(x = "Consumption (kg/person/year)") + 
  facet_wrap(~country) +
  theme_classic()
```

To answer the grand research question:

```{r}
#| fig-height: 22
#| fig-width: 11

ggplot(fc, aes(x = consumption, fill = country)) + 
  geom_density() +
  labs(x = "Consumption (kg/person/year)") + 
  facet_wrap(~food_category) +
  scale_fill_viridis_d()
```


```{r}
#| fig-height: 22
#| fig-width: 11

ggplot(fc, aes(x = food_category, y = consumption)) + 
  geom_col() + 
  facet_wrap(~country)
```





