---
title: "Bivariate Viz"
format:
  html:
    toc: true
    toc-depth: 1
    embed-resources: true
---

Use this file for practice with the **bivariate viz** in-class activity.  Refer to the class website for details.


# Examples: To Prepare for Class
```{r}
# Import data
survey <- read.csv("https://ajohns24.github.io/data/112/about_us_2024.csv")

# How many students have now filled out the survey?

nrow(survey)

# What type of variables do we have?
str(survey)
```

```{r}
# Attach a package needed to use the ggplot function
library(tidyverse)


# Make a ggplot
ggplot(survey, aes(x = hangout)) +
  geom_bar(fill = "blue") + 
  theme_minimal()

```
```{r}
#To understand temperature

ggplot(survey, aes(x = fave_temp)) +
  geom_bar(fill = "blue") + 
  theme_minimal()

ggplot(survey, aes(x = fave_temp)) +
  geom_histogram(color = "white", fill = "blue") +
  theme_minimal()
```

# In Class Exercises


```{r}
# Load data
elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")

# Check it out
head(elections)
```

# Exercise 0: Review

Part a.
I would guess 65%?

```{r}
ggplot(elections, aes(x = winner_20)) +
  geom_bar(fill = "blue") + 
  theme_minimal()

```

Part b.
```{r}
ggplot(elections, aes(x = repub_pct_20)) +
  geom_histogram(fill = "blue") + 
  theme_minimal()

```

# Exercise 1: Quantitative vs Quantitative Intution Check

```{r}
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point() + 
  geom_smooth() + 
  theme_minimal()
```
# Exercise 2: 2 Quantitiative Variables
```{r}
# Set up the plotting frame
# How does this differ than the frame for our histogram of repub_pct_20 alone? We have two different axis of different variables
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16))
```

```{r}
# Add a layer of points for each county
# Take note of the geom! --> Geom_point!Cause there's a point for each
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point()
```

```{r}
# Change the shape of the points
# What happens if you change the shape to another number? Different point shapes!
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point(shape = 4)
```

```{r}
# YOU TRY: Modify the code to make the points "orange"
# NOTE: Try to anticipate if "color" or "fill" will be useful here. Then try both. color worked!
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point(color = "orange" )
```

```{r}
# Add a layer that represents each county by the state it's in
# Take note of the geom and the info it needs to run! geom_text instead of geom_point, needs an additional variable to generate it
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_text(aes(label = state_abbr))
```

# Exercise 3: Reflect
Texas seemed to be some of the biggest outliers, however the relationship seems to be very strong. It appears to be a positive relationship. 

# Exercise 4: Visualizing trend
```{r}
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point() +
  geom_smooth()
```
Part a.
```{r}
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_smooth()
```
Part b.
```{r}
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point() +
  geom_smooth(method = "lm")
```
# Exercise 5: Your Turn
```{r}
# Scatterplot of repub_pct_20 vs median_rent
ggplot(elections, aes(y = repub_pct_20, x = median_rent)) +
  geom_point() +
  geom_smooth(method = "lm")


# Scatterplot of repub_pct_20 vs median_age
ggplot(elections, aes(y = repub_pct_20, x = median_age)) +
  geom_point() +
  geom_smooth(method = "lm")
```
From these two scatter plots, it appears that the median_rent has a stronger relationship than median_age as the linear relationship is a bit stronger. The relationship is negative between median_rent, as rent is lower higher republican % is higher.

# Exercise 6: A Sad Scatterplot

A scatter plot would be challenging comparing a quantitative variable and a categorical variable, these would be a challenging numeric variable relationships. 

```{r}
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_point()
```

# Exercise 7: Quantitative vs Categorical - Violins & Boxes

```{r}
# Side-by-side violin plots
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_violin()
```
```{r}
# Side-by-side boxplots (defined below)
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_boxplot()
```
Reflect: It appears that county-level support was higher on average in the red states, and lower in blue states but was still higher than 50% for all. 

# Exercise 8: Quantitative vs Categorical - Intuition Check
```{r}
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density()

```


# Exercise 9: Quantitative vs Categorical - Density Plots
```{r}
# Name two "bad" things about this plot It is hard to understand what is really happening. It is also hard to understand the relative scales. 

ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density()
```

```{r}
# What does scale_fill_manual do? Sets the colors for each density plot

ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density() +
  scale_fill_manual(values = c("blue", "purple", "red"))
```

```{r}
# What does alpha = 0.5 do? This changes the transparency values

# Play around with different values of alpha, between 0 and 1
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density(alpha = 0.75) +
  scale_fill_manual(values = c("blue", "purple", "red"))
```

```{r}
# What does facet_wrap do?! Sets each up on its own graph path by category

ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density() +
  scale_fill_manual(values = c("blue", "purple", "red")) +
  facet_wrap(~ historical)
```

```{r}
# Let's try a similar grouping strategy with a histogram instead of density plot.
# Why is this terrible? There is just so much going on and it is all on top of eachtother.

ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_histogram(color = "white") +
  scale_fill_manual(values = c("blue", "purple", "red"))
```
# Exercise 10
Personally, I am a fan of density plots separated by category for visualizing the relationships between quantitative and categorical variables. Density plots allow for variation relative to box plots show a little more dynamics of the relationships relative to eachother. Density plots however do not as easily show the average or the quartiles as easily as boxplots do. 

# Exercise 11: Categorical vs Categorical - Intuition Check
```{r}
# Plot 1: adjust this to recreate the top plot
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar()
```
```{r}
# Plot 2: adjust this to recreate the bottom plot
ggplot(elections, aes(x = winner_20)) +
  geom_bar() + 
  facet_wrap (~ historical)
```
# Exercise 12: Categorical vs Categorical
```{r}
# A stacked bar plot
# How are the "historical" and "winner_20" variables mapped to the plot, i.e. what roles do they play? Historical is the x axis - the predictor, while winner_20 is the response variable. 

ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar()
```

```{r}
# A faceted bar plot
ggplot(elections, aes(x = winner_20)) +
  geom_bar() +
  facet_wrap(~ historical)
```

```{r}
# A side-by-side bar plot
# Note the new argument to geom_bar sets them separated under together

ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "dodge")
```

```{r}
# A proportional bar plot
# Note the new argument to geom_bar
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "fill")
```

