---
title: "Multivariate Viz"
format:
  html:
    toc: true
    toc-depth: 1
    embed-resources: true
---

Use this file for practice with the **multivariate viz** in-class activity.  Refer to the class website for details.

```{r}
library(tidyverse)

# Import data
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv") |> 
  mutate(date = as.Date(date))  

# Check out the first 6 rows
# What are the units of observation?
head(weather)


# How many data points do we have?
dim(weather)
nrow(weather)


# What type of variables do we have?
str(weather)
```
```{r}
ggplot(weather, aes(x = temp3pm)) +
  geom_histogram(color = "white")
```

```{r}
# Plot 1 (no facets & starting from a density plot of temp3pm)
ggplot(weather, aes(x = temp3pm)) + 
  geom_density()
```

```{r}
# Plot 2 (no facets or densities)
ggplot(weather, aes(x = temp3pm, fill = location)) +
  geom_histogram(color = "white") 
```

```{r}
# Plot 3 (facets)
ggplot(weather, aes(x = temp3pm)) +
  geom_density() + 
  facet_wrap(~ location)
```

```{r}
# Don't worry about the syntax (we'll learn it soon)
woll <- weather |>
  filter(location == "Wollongong") |> 
  mutate(date = as.Date(date))  
```

```{r}
# How often does it raintoday?
# Fill your geometric layer with the color blue.

```

```{r}
# If it does raintoday, what does this tell us about raintomorrow?
# Use your intuition first
ggplot(woll, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar()
```

```{r}
# Side-by-side bars
ggplot(woll, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar(position = "dodge")

# Proportional bars
# position = "fill" refers to filling the frame, nothing to do with the color-related fill
ggplot(woll, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar(position = "fill")
```

```{r}
# Plot temp3pm vs temp9am
# Change the code in order to indicate the location to which each data point corresponds
ggplot(weather, aes(y = temp3pm, x = temp9am)) + 
  geom_text(aes(label = location)) 
```
```{r}
# Change the code in order to indicate the location to which each data point corresponds
# AND identify the days on which it rained / didn't raintoday
ggplot(weather, aes(y = temp3pm, x = temp9am, color = location, shape = raintoday)) + 
  geom_point() 
```

```{r}
# Change the code in order to construct a line plot of temp3pm vs date for each separate location (no points!)
ggplot(weather, aes(y = temp3pm, x = date)) + 
  geom_line() + 
  facet_wrap(~location)
  
```

```{r}
# Plot the relationship of raintomorrow & raintoday
# Change the code in order to indicate this relationship by location
ggplot(weather, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar(position = "fill") + 
  facet_wrap(~location)
```

# In Class Exercises

```{r}
# Import and check out data
education <- read.csv("https://mac-stat.github.io/data/sat.csv")
head(education)
```
# Exercise 1: SAT scores

```{r}
ggplot(education, aes(x = sat)) +
  geom_histogram(color = "white") 


ggplot(education, aes(x = sat)) + 
  geom_density()
```

The distribution of SAT scores appears a bit bimodal, operating on the range of potential scores from a little under 800 to close to 1200. The peak appears at around 900, then there is a bit of a dip and then another set of peaks around 1050.


# Exercise 2:
```{r}
# Construct a plot of sat vs expend
# Include a "best fit linear regression model" (HINT: method = "lm")
ggplot(education, aes(y = sat, x = expend)) + 
  geom_point() + 
  geom_smooth(method = "lm")


```


```{r}
# Construct a plot of sat vs salary
# Include a "best fit linear regression model" (HINT: method = "lm")

ggplot(education, aes(y = sat, x = salary)) + 
  geom_point() + 
  geom_smooth(method = "lm")

```

These relationships are both very interesting, as they are negative. As salary and state expenditure both increase, the negative trend shows that SAT scores fall. 


# Exercise 3:

```{r}
ggplot(education, aes(y = sat, x = salary, color = expend)) + 
  geom_point (alpha = 0.75) + 
  geom_smooth(method = "lm")

```

# Exercise 4:
```{r}
ggplot(education, aes(y = sat, x = salary, color = cut(expend, 3))) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm")
```

It appears that as salary increases for the lowest subsection of expendature sat scores see a negative relationship that is very steep.Then as salary increases and expend is in the middle group there is still a negative relationship seen with SAT scores but it is weaker. Lastly, the highest level of salaries sees the highest levels of expendature and there is a weak positive relationship with SAT scores. 

# Exercise 5:

Part a
```{r}
ggplot(education, aes(x = fracCat)) + 
  geom_bar() 

```


Part b
```{r}
ggplot(education, aes(x = sat, fill = fracCat)) +
  geom_density( alpha = 0.5) + 
  theme_minimal()

ggplot(education, aes(y = sat, x = fracCat)) +
  geom_boxplot()

```


Part c
```{r}
ggplot(education, aes(y = sat, x = expend, color = fracCat)) +
  geom_point(alpha = 0.75) +
  geom_smooth (method = lm)

```

This does not suggest that SAT scores decrease as spending increases and actually suggests the opposite. We see that for each of the three groups of states, as their spending increases so too do their SAT scores.

Part d
Maybe because there are three such distinct groups and the relationship heavily depends on this variable. For expendatures it matters per student. With the addition of these other variables we are able to separate data points out which allows for the true relationships to be found. 


North Carolina appears to be an outlier in regards to the ratio variable. Arkansas and Mississippi seem to be similar and so do New Jersey and Maryland. States that are geographically similar appear to be similar in education trends. Putting this all together you can see what the overall trends are across the US and where some of the biggest differences lie. 
