---
title: "Spatial Viz"
---

Use this file for practice with the **spatial viz** in-class activity.  Refer to the class website for details.

```{r}
# Load tidyverse package for plotting and wrangling
library(tidyverse)

# Import the data and only keep 2 addresses
energy <- read.csv("https://mac-stat.github.io/data/MacNaturalGas.csv") |> 
  mutate(date = as.Date(paste0(month, "/1/", year), "%m/%d/%Y")) |> 
  filter(address != "c")

# Check it out
head(energy)
```

```{r}
#Example 1
ggplot(energy, aes(x = therms, color = address)) + 
  geom_density()
```


```{r}
#Example 2
ggplot(energy, aes(y = therms, x = date, color = address)) + 
  geom_point() 
```

```{r}
#Example 3
ggplot(energy, aes(x = therms, fill = renovated)) + 
  geom_density(alpha = 0.5) + 
  facet_wrap(~ address)
```

```{r}
ggplot(energy, aes(y = therms, x = hdd, color = renovated)) + 
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~ address)
```

# Exercise 1
```{r}
fave_places <- read.csv("https://hash-mac.github.io/stat112site-s25/data/our_fave_places.csv")

# Check it out
head(fave_places)
```
```{r}
# Load the leaflet package
library(leaflet)

# Just a plotting frame
leaflet(data = fave_places)
```

```{r}
# Now what do we have?
leaflet(data = fave_places) |> 
  addTiles()
```

```{r}
# Now what do we have?
# longitude and latitude refer to the variables in our data
leaflet(data = fave_places) |> 
  addTiles() |> 
  addMarkers(lng = ~longitude, lat = ~latitude)
```

```{r}
# Since we named them "longitude" and "latitude", the function
# automatically recognizes these variables. No need to write them!
leaflet(data = fave_places) |> 
  addTiles() |> 
  addMarkers()
```

```{r}
# Load package needed to change color
library(gplots)

# We can add colored circles instead of markers at each location
leaflet(data = fave_places) |> 
  addTiles() |> 
  addCircles(color = col2hex("red"))
```

```{r}
# We can change the background
# Mark locations with yellow dots
# And connect the dots, in their order in the dataset, with green lines
# (These green lines don't mean anything here, but would if this were somebody's travel path!)
leaflet(data = fave_places) |>
  addProviderTiles("USGS") |>
  addCircles(weight = 10, opacity = 1, color = col2hex("yellow")) |>
  addPolylines(
    lng = ~longitude,
    lat = ~latitude,
    color = col2hex("green")
  )
```

```{r}
# Import starbucks location data
starbucks <- read.csv("https://mac-stat.github.io/data/starbucks.csv")
```

```{r}
# Don't worry about the syntax
starbucks_mn <- starbucks |>   
  filter(Country == "US", State.Province == "MN")
```


# Part 2

```{r}
# Just a plotting frame
leaflet(data = starbucks_mn) |> 
  addTiles() |> 
  addCircles(color = col2hex("red"),lng = ~Longitude, lat = ~Latitude)
  

```

```{r}
ggplot(starbucks, aes(x = Longitude, y = Latitude)) + 
  geom_point(alpha = 0.2, size = 0.2)

```

This is not super interactive but it is cool to see the outline of the world. 

```{r}
# Load the package
library(rnaturalearth)

# Get info about country boundaries across the world
# in a "sf" or simple feature format
world_boundaries <- ne_countries(returnclass = "sf")
```

```{r}
# What does this code produce? a map of the outline of the countries of the world

# What geom are we using for the point map? geom_sf
ggplot(world_boundaries) + 
  geom_sf()
```

```{r}
# Load package needed to change map theme
library(mosaic)

# Add a point for each Starbucks
# NOTE: The Starbucks info is in our starbucks data, not world_boundaries
# How does this change how we use geom_point?!
ggplot(world_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3, size = 0.2, color = "darkgreen"
  )+
  theme_map()
```
They are heavily concentrated in certain countries, also along the coasts. There are so many in the US.

```{r}
# We'll learn this syntax soon! Don't worry about it now.
starbucks_cma <- starbucks |> 
  filter(Country %in% c('CA', 'MX', 'US'))
```

```{r}
cma_boundaries <- ne_states(
  country = c("canada", "mexico", "united states of america"),
  returnclass = "sf")
```

```{r}
# Just the boundaries
ggplot(cma_boundaries) + 
  geom_sf()
```

```{r}
# Add the points
# And zoom in
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3,
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50)) +
  theme_map()
```

```{r}
starbucks_midwest <- starbucks |> 
  filter(State.Province %in% c("MN", "ND", "SD", "WI"))
```

```{r}
# Load packages
library(sf)
library(maps)

# Get the boundaries
midwest_boundaries <- st_as_sf(
  maps::map("county",
            region = c("minnesota", "wisconsin", "north dakota", "south dakota"), 
            fill = TRUE, plot = FALSE))

# Check it out
head(midwest_boundaries)
```

```{r}
ggplot(midwest_boundaries) + 
    geom_sf() + 
    geom_point(
    data = starbucks_midwest,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3,
    size = 0.2, 
    color = 'darkgreen'
    ) + 
    theme_map()
```

#Exercise 7

```{r}
# Point map (we made this earlier)
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3,
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50), ylim = c(14.54, 83.11)) +
  theme_map()
```

```{r}
# What changed in the plot?
# What changed in our code?!
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_density_2d(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    linewidth = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50), ylim = c(14.54, 83.11)) +
  theme_map()
```

#Part 3: Choropleth Maps

```{r}
elections_by_state <-  read.csv("https://mac-stat.github.io/data/election_2020_by_state.csv")
elections_by_counties <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")
```

```{r}
# Don't worry about the code!

elections_by_state <- elections_by_state |> 
  filter(state_abbr != "DC") |> 
  select(state_name, state_abbr, repub_pct_20) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(30, 70, by = 5), 
               labels = c("30-34", "35-39", "40-44", "45-49",
                          "50-54", "55-59", "60-64", "65-70"), 
               include.lowest = TRUE))

elections_by_counties <- elections_by_counties |> 
  select(state_name, state_abbr, county_name, county_fips,
          repub_pct_20, median_age, median_rent) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(0, 100, by = 10),
               labels = c("0-9", "10-19", "20-29", "30-39", "40-49",
                          "50-59", "60-69", "70-79", "80-89", "90-100"),
               include.lowest = TRUE))
```

```{r}
# Get the latitude and longitude coordinates of state boundaries
states_map <- map_data("state")

# Check it out
head(states_map)
```

```{r}
head(states_map)
```

```{r}
head(elections_by_state) 
```

```{r}
# Note where the dataset, elections_by_state, is used
# Note where the background map, states_map, is used
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_pct_20)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() 
```

```{r}
# Make it nicer!
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_pct_20)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_gradientn(name = "% Republican", colors = c("blue", "purple", "red"), values = scales::rescale(seq(0, 100, by = 5)))
```

```{r}
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map()
```

```{r}
# Load package needed for refining color palette
library(RColorBrewer)

# Now fix the colors
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_manual(values = rev(brewer.pal(8, "RdBu")), name = "% Republican")
# Load package needed for refining color palette
```

```{r}
# Get only the starbucks data from the US
starbucks_us <- starbucks |> 
  filter(Country == "US")

# Map it
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  geom_point(
    data = starbucks_us,
    aes(x = Longitude, y = Latitude),
    size = 0.05,
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_manual(values = rev(brewer.pal(8, "RdBu")), name = "% Republican")
```

This is geom_map. 


```{r}
# Get the latitude and longitude coordinates of county boundaries
library(socviz)
data(county_map) 

# Check it out
head(county_map)
```

```{r}
head(county_map)

```

```{r}
head(elections_by_counties)
```
```{r}
# Add 0's at the beginning of any fips_code that's fewer than 5 numbers long
# Don't worry about the syntax
elections_by_counties <- elections_by_counties |> 
  mutate(county_fips = as.character(county_fips)) |> 
  mutate(county_fips = 
           ifelse(nchar(county_fips) == 4, paste0("0", county_fips), county_fips))
```

```{r}
ggplot(elections_by_counties, aes(map_id = county_fips, fill = repub_20_categories)) +
  geom_map(map = county_map) +
  scale_fill_manual(values = rev(brewer.pal(10, "RdBu")), name = "% Republican") +
  expand_limits(x = county_map$long, y = county_map$lat) +
  theme_map() +
  theme(legend.position = "right") + 
  coord_equal()
```

```{r}
ggplot(elections_by_counties, aes(map_id = county_fips, fill = median_age)) +
  geom_map(map = county_map) +
  expand_limits(x = county_map$long, y = county_map$lat) +
  theme_map() +
  scale_fill_gradientn(colors = c("blue", "purple", "red"), values = scales::rescale(seq(0, 100, by = 5))) 

```

```{r}
ggplot(elections_by_counties, aes(map_id = county_fips, fill = median_rent)) +
  geom_map(map = county_map) +
  expand_limits(x = county_map$long, y = county_map$lat) +
  theme_map() +
  scale_fill_gradientn(colors = c("blue", "purple", "red"), values = scales::rescale(seq(0, 100, by = 5))) 

```
# Exercise 11: Choropleth maps with Leaflet

```{r}
states <- sf::read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson")
class(states)
names(states)
```

```{r}
m <- leaflet(states) %>%
  setView(-96, 37.8, 4) %>%
  addTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
```

```{r}
m %>% addPolygons()
```

```{r}
bins <- c(0, 10, 20, 50, 100, 200, 500, 1000, Inf)
pal <- colorBin("YlOrRd", domain = states$density, bins = bins)

m %>% addPolygons(
  fillColor = ~pal(density),
  weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7)
```

```{r}
m %>% addPolygons(
  fillColor = ~pal(density),
  weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7,
  highlightOptions = highlightOptions(
    weight = 5,
    color = "#666",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE))
```

```{r}
labels <- sprintf(
  "<strong>%s</strong><br/>%g people / mi<sup>2</sup>",
  states$name, states$density
) %>% lapply(htmltools::HTML)

m <- m %>% addPolygons(
  fillColor = ~pal(density),
  weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7,
  highlightOptions = highlightOptions(
    weight = 5,
    color = "#666",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE),
  label = labels,
  labelOptions = labelOptions(
    style = list("font-weight" = "normal", padding = "3px 8px"),
    textsize = "15px",
    direction = "auto"))
m
```

```{r}
m %>% addLegend(pal = pal, values = ~density, opacity = 0.7, title = NULL,
  position = "bottomright")
```


```{r}
# From https://leafletjs.com/examples/choropleth/us-states.js
states <- sf::read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson")

bins <- c(0, 10, 20, 50, 100, 200, 500, 1000, Inf)
pal <- colorBin("YlOrRd", domain = states$density, bins = bins)

labels <- sprintf(
  "<strong>%s</strong><br/>%g people / mi<sup>2</sup>",
  states$name, states$density
) %>% lapply(htmltools::HTML)

leaflet(states) %>%
  setView(-96, 37.8, 4) %>%
  addTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN'))) %>%
  addPolygons(
    fillColor = ~pal(density),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>%
  addLegend(pal = pal, values = ~density, opacity = 0.7, title = NULL,
    position = "bottomright")
```

